---
title: "Animal_Hierarchal_Clustering"
author: "Natasha Santhanam"
date: "4/5/2021"
output: html_document
---

## Here we look at different methods for Cluster Analysis of Reaction IDs in our phenotype matrix

```{r}
library(tidyverse)
library(workflowr)
library(glue)
library(factoextra)
library(NbClust)
library(cluster)    
library(ggfortify)
"%&%" = function(a,b) paste(a,b,sep="")
dir <- "/Users/natashasanthanam/Github/rat-genomic-analysis/data/"

all_animals <- read_tsv(dir %&% "all_animals_h2.txt", col_names = TRUE)
```



First I'll use the Factoextra package in R to determine what is the appropriate number of clusters. 

```{r}
cluster_matrix <- all_animals[,-1]
cluster_matrix <- scale(cluster_matrix)

#elbow method
fviz_nbclust(cluster_matrix, kmeans, method = "wss") +
    geom_vline(xintercept = 4, linetype = 2)+
  labs(subtitle = "Elbow method")

#It'll definitely be 4 since we have 4 species but still good to check
```


After determining the best k, we can create a new fit
```{r}
fit <- kmeans(cluster_matrix, 4) # 4 cluster solution
# get cluster means 
aggregate(cluster_matrix,by=list(fit$cluster),FUN=mean)
# append cluster assignment
cluster_matrix_k <- data.frame(cluster_matrix, fit$cluster)

k2 <- kmeans(cluster_matrix, centers = 4, nstart = 25)
str(k2)
fviz_cluster(k2, data = cluster_matrix)

```

Hierarchical Agglomerative # Ward Hierarchical Clustering
```{r}
d <- dist(cluster_matrix, method = "euclidean") # distance matrix
fit <- hclust(d, method="ward") 
plot(fit) # display dendogram
groups <- cutree(fit, k=4) # cut tree into 4clusters
# draw dendogram with red borders around the 4 clusters 
rect.hclust(fit, k=4, border="red")
```

Visualize Distance Matrix between Reaction IDs
```{r}

fviz_cluster(list(data = cluster_matrix_k, cluster = cluster_matrix_k$fit.cluster), palette = "Set2", ggtheme = theme_minimal())

```
DO PCA Component Analysis with all species
```{r}
h2_matrix <- as.matrix(cbind(species = c(1, 2, 3, 4),h2_matrix))

pca_res <- prcomp(h2_matrix, scale. = TRUE)
autoplot(pca_res, data = h2_matrix, colour = 'species', label = TRUE, label.size = 0.5)
```

