---
title: "rat_MultiXcan"
author: "Natasha Santhanam"
date: "10/18/2021"
output: html_document
---

```{r setup, include=FALSE}
suppressMessages(library(tidyverse))
suppressMessages(library(data.table))
suppressMessages(library(RSQLite))
suppressMessages(library(qqman))
"%&%" = function(a,b) paste(a,b,sep="")
```

## Run MultiXcan across all 5 Rat Tissues and use Metabolic Traits as phenotypes

# Check correlation of observed expression and predicted expression (from prediXcan of metabolic traits)

```{r corr between obs expr and metabolic traits, eval=FALSE}
expr <- as.data.frame(readRDS("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/expression/Ac_expression_transformed.RDS"))
pred_expr <- read_tsv("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan/rat_metabolic_Ac__predict.txt", col_names = TRUE)

expr <- expr %>% mutate(IID = rownames(expr), .before = "ENSRNOG00000000007")
pred_expr <- pred_expr[na.omit(match(expr$IID, pred_expr$IID)), na.omit(match(colnames(expr), colnames(pred_expr))) ]
expr <- expr[na.omit(match(pred_expr$IID, expr$IID)), na.omit(match(colnames(pred_expr), colnames(expr))) ]

corr_both <- cor(pred_expr, expr)
corr_diagonal <- diag(corr_both)
hist(corr_diagonal)
```


```{r plot correlations against R2 from models}
filename <- "/Users/natashasanthanam/Box/imlab-data/data-Github/rat-genomic-analysis/sql/Ac_output_db.db"
sqlite.driver <- dbDriver("SQLite")
  conn <- dbConnect(RSQLite::SQLite(), filename)
  pred_r2 <- dbGetQuery(conn, 'select * from extra')

corr_diagonal <- as.data.frame(corr_diagonal) 
corr_diagonal <- cbind(gene = rownames(corr_diagonal), corr_diagonal)

all_cor_r2 <- inner_join(corr_diagonal, pred_r2, by = "gene")
ggplot(all_cor_r2, aes(corr_diagonal, R2)) + geom_point() + xlab("Correlation between pred expr vs obs expr") + ylab ("cv predicted peformance from Tyson generated models")
```



# Generate Folder with Predicted Expression Data for each Tissue

First have to remove potential overlap between genos used in predicted expresssion and those in phenotype file. Should only be around 60ish so not too big a deal

```{r generate list of ids across all tissues, eval=FALSE}
filelist <- list.files("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/expression", pattern = ".RDS", full.names = TRUE)
all_names <- data.frame(ID = as.character())
for(fila in filelist) {
  tempo <- readRDS(fila)
  tempo <- as.data.frame(rownames(tempo)) %>% rename(ID = `rownames(tempo)`)
  all_names <- full_join(tempo, all_names, by = "ID")
}
```


```{r clean up pheno file, eval=FALSE}
pheno <- read_csv("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/Box_files/processed_obesity_rat_Palmer_phenotypes.csv", col_names=TRUE)

pheno <- pheno %>% rename(ID = rat_rfid) %>% filter(!ID %in% all_names$ID)
#write_tsv(pheno, "/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/metabolic_trait_phenos_MultiXcan.txt", col_names = TRUE)
```


Next have to remove overlap rats predicted expression as well
```{r  generate expr data, eval=FALSE}
filelist <- list.files("/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/prediXcan", pattern = "__predict.txt", full.names = TRUE)

for(fila in filelist) {
  tempo <- fread(fila, header=TRUE)
  name <- substr(fila, 72,73)
  tempo <- tempo %>% filter(!FID %in% all_names$ID)
  tempo <- tempo[match(pheno$ID, tempo$FID),]
  write_tsv(tempo, "/gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/expr/" %&% name %&% ".txt")
}
```


Run MultiXcan using the predicted expression from prediXcan across all 5 tissues to boost power
```{bash run MultiXcan}
#!/bin/bash

#PBS -N multixcan
#PBS -S /bin/bash
#PBS -l walltime=4:00:00
#PBS -l mem=4gb
#PBS -l nodes=1:ppn=1

# SPECIFY LOGGING BEHAVIOR

#PBS -o /gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/logs/${PBS_JOBNAME}.${PBS_JOBID}.log
#PBS -e /gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/logs/${PBS_JOBNAME}.${PBS_JOBID}.err

module load gcc/6.2.0
source ~/.bashrc 
conda activate /gpfs/data/im-lab/nas40t2/bin/envs/tensorqtl/

echo "MultiXcan running on epifat"

python /gpfs/data/im-lab/nas40t2/natasha/GTEX_Analysis/MetaXcan/software/MulTiXcan.py \
        --expression_folder /gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/expr \
        --expression_pattern "(.*)_expression_transformed.txt" \
        --input_phenos_file /gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/metabolic_trait_phenos_MultiXcan.txt \
        --input_phenos_column fasting_glucose \
        --output /gpfs/data/im-lab/nas40t2/natasha/rat_genomics/MultiXcan/results/fasting_glucose_predict_assoc.txt \
        --pc_condition_number 10 \
        --mode linear \
        --verbosity 8 \
        --throw
        
```


